### 2020-09-01 LaTeX/sections/paper.tex
Paper- Edits

\item\textbf{{Mapping with keys stored in Heap.}} This priority queue stores items inside a Solidity mapping that maps the mapping keys (256-bit integers) to these items. The mapping keys are then stored (based on the value they are mapped to) in a sorted Heap that is implemented with a dynamically-sized array.

@@

Then, we called the \texttt{Dequeue()} function which iteratively removes the highest priority number from the priority queues (until the data structure is empty). The computational costs for dequeuing 50 unsigned integers in each priority queue are outlined in Table~\ref{tab:PQUnitTests}. These tests are performed using the current Ethereum gas metrics (block gas limit $=$11,741,495 and 1 gas $=$ 56 gwei)~\footnote{https://ethstats.net/}. The second column of the table shows the net gas consumption (the gasUsed value derived from transaction receipts) for removing 50 integers from each priority queues.

@@

@@

%& \textbf{\thead{Net Cost\\in Gas\\ (\texttt{gasUsed})}}  
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  %
% A little explanation on gas metrics (Mahsa: please let me know if this is still unclear:)

%First let's take a look at the following example:

%Suppose we have a smart contract which uses up 14,000 gas. The gas limit that we have set up is 20,000 gas. The smart contract also includes an SSTORAGE command.
%So, how much gas will the contract creator get back post computation?
%Firstly, they will get back (20,000-14,000) = 6,000 units of unused gas.
%Now, the SSTORAGE command has also been used, so theoretically they should get back 15,000 gas as well.
%However, the amount of gas that has been used in the contract is 14,000 and since 15,000 > 14,000/2, the REFUND generated will be 14,000/2= 7000.
%So the total gas that the creator is getting back in the end is 6000+7000 = 13,000.

@@

Based on our tests results, the \textit{linked list} and \textit{linked list with mapping} priority queues show the best performance in compared to the different heap implementations. In the next section, we explore the differences between the two different implementations of the linked list priority queue.

@@

\subsection{Linked List \textit{vs.} Linked List with Mapping}

There are two main ways to implement each of the \textit{linked list} and \textit{linked list with mapping} priority queues. In this section, we perform a comprehensive comparison between these two with the purpose choosing the proper data structure for implementing the call market.

@@

@@

% = = = = = = = = = = = = = = = = = = = = =Call market operations Table = = = = = = = = = = = =  = = = = = =  %
\begin{table}[]
\begin{tabular}{|l|l|}
\hline
\multicolumn{1}{|c|}{\textbf{Operation}} & \multicolumn{1}{c|}{\textbf{Description}}                            			\\ \hline

@@

Table~\ref{tab:worst_case_matching} outlines the \texttt{Match()} function's computational costs and performance for three different implementations of the call market. We used three different designs of the linked list with mapping priority queue. (1) Every time a trade happens, we use the \texttt{DELETE} method to remove the mapping elements that contains best bid and ask orders from storage completely \textbf{and} update the head and tail pointers of the list. (2) We only update the pointers of the priority queue and do not remove the best bid and ask orders during the entire matching process. Instead we destroy the priority queue smart contract~\footnote{We execute a public function defined in the priority queue smart contract which itself calls the \texttt{self-destruct()} operation and transfers its funds (if any) to the call market contract.} once the matching is completed. (3) In the last implementation, we allow the matching process to be completed and do not clear any on chain data. As it can be seen in Table~\ref{tab:worst_case_matching}, the \texttt{Match()} operation shows a better performance when killing the priority contract and/or not clearing data from the chain. Although deleting the priority queue contract adds an extra external call to the \texttt{Match()} function and should technically increase the gas cost, the 24000 gas refunds from \texttt{SELFDESTRUCT} would compensate the cost. So it can still match as same number of orders (before running out of gas)as in the no on chain clearing implementation. The last column of Table~\ref{tab:worst_case_matching} reflects the gas cost for matching 1000 pairs of bids and asks for which we set the block gas limit to the maximum of  $2^{53}$ (the Javascript's max safe integer).

@@

@@

%  = = = = = = = = = = % 
%Mahsa: Im not sure which of these tables are good for this section!?
%  = = = = = = = = = = %

@@

To maintain the collateral balance of each market participant, we use two Solidity type one-to-one mapping that map Ethereum addresses to 256 bits unsigned integers; \texttt{TotalBalance} and \texttt{UnavailableBalance}. Once market closes and orders are matched, the \texttt{UnavailableBalance} needs to be cleared. However, since it is not possible to delete the entire mapping without knowing the keys~\footnote{https://solidity.readthedocs.io/en/v0.5.12/security-considerations.html}, clearing the \texttt{UnavailableBalance} mapping remains a challenging issue to solve . Here we provide a landscape of solutions for that.
%only individual keys and what they map to can be deleted.

%---------+++++++++++++++++++-----------------
\item \textbf{Creating a New Mapping Every Time the Market Opens.} Instead of clearing the \texttt{UnavailableBalance} of traders at the end of a matching process, we could create a new mapping every time the market opens. Note that using this solution, traders can only claim their funds (using the \texttt{ClaimEther()} and/or \texttt{ClaimToken} methods) only when the market is in state \texttt{Closed}.

%---------+++++++++++++++++++-----------------
\item \textbf{Creating Custom Keys for the Mapping.} We can create custom keys for the mapping by defining a counter as a global variable inside the CallMarket smart contract. This counter is incremented at the end of the matching process. So instead of clearing the mapping, we only use another portion of it every time the market opens.

%---------+++++++++++++++++++-----------------
\item \textbf{Storing the Mapping in a Data-Contract.} Another design proposal is to create a smart contract every time the market opens, this data contract only stores the \texttt{UnavailableBalance} mapping and will be killed at the end of the matching process.

%---------+++++++++++++++++++-----------------
\item \textbf{Storing the Mapping keys in an additional Array.} Another common pattern is to create an additional array on top of the mapping and iterate over that. This array (\eg address[]) stores the traders' addresses and enables us to iterate over the mapping and delete individual keys and what they map to at the end of the matching process. Not that this is a gas-costly design pattern as we would need to maintain a secondary data structure.

% = = = = = = = = = = Design Details of PQs = = = = = = = = = = = = = = %
\section{Design Details of PQs}

\subsection{Mapping with Keys Stored in a Heap}\label{sec:mapping_heap}

To store the orders, here we use Solidity mapping that map orders' ids (256-bit integers) to order structs with their variables of different types. The mapping keys are then stored in a sorted heap that is implemented with a dynamic storage array. Every time a trade happens; (i) the mapping keys associated with the best bid and ask orders are deleted from the heap and the heap is re-sorted, and (ii) the mapping elements containing the best bid and ask orders are deleted from storage completely which refund 15000 gas~\cite{wood2014ethereum}. Deleting the mapping elements is yet a gas-costly operation which lowers down the maximum number of orders the \texttt{Match()} function could handle in this case.~\footnote{Every order structs contains multiple variables of different types and removing them from storage is identical to setting multiple variables to zero.} However, we clear the mapping element once a trade happens in our design as leaving unnecessary data is not a proper design practice.

%---------+++++++++++++++++++-----------------
\subsection {Linkedlist}
Every time a trade happens, we use the \texttt{SELFDESTRUCT} operation to remove the nodes (\ie smart contracts) containing the best bid and ask orders from the Ethereum blockchain. When a node is deleted, it transfers its funds (if any) to the payable address of the CallMarket contract that has been previously passed to it as a constructor argument. Also, removing a smart contract from the Ethereum blockchain refunds 24000 gas to the caller.

%---------+++++++++++++++++++-----------------
\subsection {Linkedlist with Mapping}
Every time a trade happens when the \texttt{Match()} function is executed, the best bid and ask orders need to be removed from the data structures. We could do this in two different ways; (i) update the head and tail pointers of the linkedlist and/or (ii) removing the mapping elements that contain the best bid and ask orders from storage completely as well as updating the pointers of the linkedlist. \par
For the same reason we mentioned in~\ref{sec:mapping_heap}, we clear the best bid and ask order structs from the mapping every time a trade happens as well as updating the pointers.

@@

% = = = = = = = = = = = =How to enforce time on on the Ethereum blockchain? = = = = = = = = = = = = =  %

\subsection{Enforcing Time on the Blockchain}

@@

\subsection{Who Pays the Cost for Closing and Reopening the Market?}

@@

%= = = = = = Miners Close the Market = = = = = = =%

\item \textbf {Miners Close and Reopen the Market.} The difference between the best bid and ask prices is called the \textit{bid-ask spread}. In our design, when the trade occurs between the the highest bid (the highest amount a buyer is willing to pay for an asset) and lowest ask (the lowest amount a seller is willing to accept for an asset) orders, the bid-ask spread is paid to the miner. There are possibilities that (i) no trade occurs or (ii) the bid-ask spread is zero (\ie the best bid and best ask prices are identical). So there is enough economic incentive for the the miner to execute the \texttt{CloseMarket()} function and get refunded as the refund amount could be potentially higher than the bid-ask spread.

% = = = = = = = = = =Processing Orders in Groups= = = = = = = = = = = = = = = %

\item \textbf{Processing Orders in Groups.} Another solution pattern is to process certain number of bid and ask orders upon every execution of the \texttt{CloseMarket()} function rather than treating them as one substantial transaction. A market participant $P_i$ would process $n$ orders from the previous market (\texttt{CloseMarket(n)}) when sending new orders to the current market. This process continues until all the orders in the previous markets are processed. % incentive? => contribution to the community!?

%= = = = = =  = =  =  Meta txs =  = = = = = = = = =%

@@

>>>>

\item\textbf{{Mapping with Keys Stored in Heap.}} This priority queue stores items inside a Solidity mapping that maps the mapping keys (256-bit integers) to these items. The mapping keys are then stored (based on the value they are mapped to) in a sorted Heap that is implemented with a dynamically-sized array.

@@

Then, we called the \texttt{Dequeue()} function which iteratively removes the highest priority number from the priority queues (until the data structure is empty). The computational costs for dequeuing 50 unsigned integers in each priority queue are outlined in Table~\ref{tab:PQUnitTests} where (\full) indicates the full refund and (\prt) means the amount of refund is bounded. These tests are performed using the current Ethereum gas metrics (block gas limit $=$11,741,495 and 1 gas $=$ 56 gwei)~\footnote{https://ethstats.net/}. The second column of the table shows the net gas consumption (the gasUsed value derived from transaction receipts) for removing 50 integers from each priority queues.

%Truffle gas limit for the tests: 11741495

@@

% Below is the same table but with "Full, Partial,None" for refunds

\textbf{\thead{Priority Queue}}    & \textbf{\thead{Net Cost in Gas\\ (\texttt{gasUsed})}}             & \textbf{\thead{Gas Refunds from Deleting a Storage}} \\ \hline

% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
	\textbf{\thead{Heap with \\ Dynamic Array}}         				& 2,518,131            		      & \full                      \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
	\textbf{\thead{Heap with \\ Static Array}}           				& 1,385,307                             & \full                      \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
	\textbf{\thead{Mapping with \\ keys stored \\ in Heap}} 		& 2,781,684                            & \full                     \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
	\textbf{Linked List}                       							& 557,085               	           & \full                      \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
	\textbf{\thead{Linked List with \\ Mapping}}          				& 731,514              	     	  &  \full                       \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %

@@

%& \textbf{\thead{Net Cost\\in Gas\\ (\texttt{gasUsed})}}  
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  %

@@

Based on our tests results, the \textit{linked list} and \textit{linked list with mapping} priority queues show a better performance in compared to the different heap implementations. In the next section, we explore the differences between the two different implementations of the linked list priority queue.

@@

\subsection{Linked List \textit{vs.} Linked List with Mapping}\label{sec:LLvsLLMapping}

There are two main ways to implement each of the \textit{linked list} and \textit{linked list with mapping} priority queues. In this section, we perform a comprehensive comparison between these two with the purpose of choosing the proper data structure for implementing the call market.

@@

% Below is the same table but with "Full, Partial,None" for refunds

\begin{table}[]
\begin{tabular}{|c|c|c|c|c|}
\hline
\multirow{2}{*}{\textbf{Gas Metrics}}                        							& \multicolumn{2}{c|}{\textbf{Linked List}}           						& \multicolumn{2}{c|}{\textbf{Linked List with Mapping}}     					\\ \cline{2-5} 
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  = = = = = =  = = = = = = = = = = = = = = = =  = = = = = =  = = = = = = %
                                                                								& with \texttt{SELFDESTRUCT} 		& without \texttt{SELFDESTRUCT}                	& with \texttt{DELETE} 		& without \texttt{DELETE}              		\\ \hline                                                                                                                                                                                                                                                          % = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  = = = = = =  = = = = = = = = = = = = = = = =  = = = = = =  = = = = = = %
                  \textbf{\thead{Net Cost in Gas\\(\texttt{gasUsed})}}                    & 557,085                    			& 721,370                     			& 731,514                     		& 334,689                      				\\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  = = = = = =  = = = = = = = = = = = = = = = =  = = = = = =  = = = = = = %
 \textbf{\thead{Gas Refund \\from Deleting a Storage}}            			& \full                 					& $\times$                 					& \full                    		& \prt                       				\\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  = = = = = =  = = = = = = = = = = = = = = = =  = = = = = =  = = = = = = %

@@

%GitHub repo:  \footnote{\CallMarketRepo}
% = = = = = = = = = = = = = = = = = = = = =Call market operations Table = = = = = = = = = = = =  = = = = = =  %
\begin{table}[]
\begin{tabular}{|l|l|}
\hline
\multicolumn{1}{|c|}{\textbf{Operation}} & \multicolumn{1}{c|}{\textbf{Description}}                            			\\ \hline

@@

Table~\ref{tab:worst_case_matching} outlines the \texttt{Match()} function's computational costs and performance for three different implementations of the call market. We used three different designs of the linked list with mapping priority queue. (1) Every time a trade happens, we use the \texttt{DELETE} method to remove the mapping elements that contains best bid and ask orders from storage completely \textbf{and} update the head and tail pointers of the list. (2) We only update the pointers of the priority queue and do not remove the best bid and ask orders during the entire matching process. Instead we destroy the priority queue smart contract~\footnote{We execute a public function defined in the priority queue smart contract which itself calls the \texttt{self-destruct()} operation and transfers its funds (if any) to the call market contract.} once the matching is completed. (3) In the last implementation, we allow the matching process to be completed and do not clear any on chain data. As it can be seen in Table~\ref{tab:worst_case_matching}, the \texttt{Match()} operation shows a better performance when killing the priority contract and/or not clearing data from the chain. However, leaving unnecessary storage data on chain is not a proper design practice (note that as mentioned in Section~\ref{sec:LLvsLLMapping}, destroying a smart contract does not remove the initial byte code of the contract from the chain but it frees up the state storage and refunds 24000 gas to the caller.)
The last column of Table~\ref{tab:worst_case_matching} reflects the gas cost for matching 1000 pairs of bids and asks for which we set the block gas limit to the maximum of  $2^{53}$ (the Javascript's max safe integer).

@@

%Although deleting the priority queue contract adds an extra external call to the \texttt{Match()} function and should technically increase the gas cost, the 24000 gas refunds from \texttt{SELFDESTRUCT} would compensate the cost. So it can still match as same number of orders (before running out of gas) as in the no on chain clearing implementation.

@@

@@

To maintain the collateral balance of each market participant, we use two Solidity type one-to-one mappings that map Ethereum addresses to 256 bits unsigned integers; \texttt{TotalBalance} and \texttt{UnavailableBalance}. Once market closes and orders are matched, the \texttt{UnavailableBalance} needs to be cleared. However, since it is not possible to delete the entire mapping without knowing the keys~\footnote{https://solidity.readthedocs.io/en/v0.5.12/security-considerations.html}, clearing the \texttt{UnavailableBalance} mapping remains a challenging issue to solve. Here we provide a landscape of solutions for that.
%only individual keys and what they map to can be deleted.

%---------+++++++++++++++++++-----------------
\item \textbf{Creating a New Mapping Every Time the Market Opens.} Instead of clearing the \texttt{UnavailableBalance} of traders at the end of a matching process, we could create a new mapping every time the market opens. Note that in this scheme traders can claim their funds back (using the \texttt{ClaimEther()} and/or \texttt{ClaimToken} methods) only when the market is closed (\ie the market is in the \texttt{Closed} state).

%---------+++++++++++++++++++-----------------
\item \textbf{Creating Custom Keys for the Mapping.} We can create custom keys for the mapping by defining a counter as a global variable inside the call market smart contract and increment it only once at the end of the matching process. So instead of clearing the mapping, we only use another portion of it every time the market opens.

%---------+++++++++++++++++++-----------------
\item \textbf{Storing the Mapping in a Data-Contract.} Another design proposal is to create a new smart contract every time the market opens, this data contract only stores the \texttt{UnavailableBalance} mapping and will be destroyed (using the \texttt{SELFDESTRUCT} operation) at the end of the matching process.

%---------+++++++++++++++++++-----------------
\item \textbf{Storing the Mapping keys in an additional Array.} Another common pattern is to create an additional array on top of the mapping and iterate over that. This array (\eg address[]) stores the traders' addresses and enables us to iterate over the mapping and delete individual keys and what they map to at the end of the matching process. Note that this is a gas-costly design pattern as we would need to maintain a secondary data structure.

@@

\begin{itemize}
% = = = = = = = = = = = =How to enforce time on on the Ethereum blockchain? = = = = = = = = = = = = =  %

\item\textbf{{Enforcing Time on the Blockchain.}}

@@

\item\textbf{{Who Pays the Cost for Closing and Reopening the Market?}}

@@

%= = = = = = Miners Close the Market = = = = = = =%
\item \textbf {Miners Close and Reopen the Market.} The difference between the best bid and ask prices is called the \textit{bid-ask spread}. In our design, when the trade occurs between the the highest bid (the highest amount a buyer is willing to pay for an asset) and lowest ask (the lowest amount a seller is willing to accept for an asset) orders, the bid-ask spread is paid to the miner. There are possibilities that (i) no trade occurs or (ii) the bid-ask spread is zero (\ie the best bid and best ask prices are identical). So there is enough economic incentive for the the miner to execute the \texttt{CloseMarket()} function and get refunded as the refund amount could be potentially higher than the bid-ask spread. 
% = = = = = = = = = =Processing Orders in Groups= = = = = = = = = = = = = = = %

\item \textbf{Processing Orders in Groups.} Another solution pattern is to process certain number of bid and ask orders upon every execution of the \texttt{CloseMarket()} function rather than treating them as one substantial transaction. A market participant $P_i$ would process $n$ orders from the previous market (\texttt{CloseMarket(n)}) when sending new orders to the current market. This process continues until all the orders in the previous markets are processed. % incentive? => contribution to the community!?
%= = = = = =  = =  =  Meta txs =  = = = = = = = = =%

@@

\end{itemize}
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  %
\end{itemize}

% = = = = = = = = = = = =Transferring the Collateral back to the traders = = = = = = = = = = = = =  %
\subsection{Transferring the Collateral back to Traders.}
